Final Project Report: Developing an AI-Assisted Grading System Using Large Language Models

### 2025-09-10 FinalProject/src/paper.tex
updated paper2

% ===================== Packages =====================
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[sc,osf]{mathpazo}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{float} % for [H] placement if desired
\usepackage{tikz} % Diagrams
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,matrix}
\usepackage{siunitx} % Optional but useful
\sisetup{detect-all}
\PassOptionsToPackage{hyphens}{url} % Make long URLs break better in refs (must come before hyperref)
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[nameinlink,capitalize,noabbrev]{cleveref}
\usepackage{bookmark} % better PDF outlines
\usepackage{csquotes} % for \enquote
\usepackage{tabularx} % better tables that auto-wrap
\urlstyle{same}
\usepackage{listings}
\usepackage{lstautogobble} % trims common indent
\usepackage{array}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{subcaption} % defines the subfigure environment
\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  breakatwhitespace=false, % Allow breaks anywhere (fixes long strings like connection URLs)
  breakindent=20pt, % Indent wrapped lines for better readability in .cs files
  postbreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\color{gray}\hookrightarrow\space}}, % Subtle continuation arrow for wrapped lines
  columns=fixed, % Changed to fixed for true monospace and even frame lines (prevents uneven extensions)
  keepspaces=true,
  showstringspaces=false,
  frame=single,
  framerule=0.5pt, % Thinner frame rule to reduce visual emphasis on any minor irregularities
  rulecolor=\color{gray}, % Gray frame for subtlety
  framesep=5pt, % Inner padding to keep content from edges
  autogobble=true,
  xleftmargin=5pt, % Slightly wider left margin to prevent hugging
  xrightmargin=5pt, % Matching right margin
  tabsize=4 % Standardize tabs (fixes potential "weird" alignment in .cs code)
}

@@

\begin{abstract}
We present an instructor-in-the-loop grading system that accelerates evaluation of open-ended student work across scanned and digital workflows. The system crops answer regions from PDFs, assigns submissions via OCR on identity regions only, and groups answers by visual semantics using a vision LLM. Instructors review and edit groups, apply rubric items once per group, and export grades from an on-screen table. The solution integrates Ghostscript rasterization, PdfPig page orchestration, SkiaSharp region extraction, Tesseract identity OCR, and GPT-4o Vision for grouping\cite{ghostscript,pdfpig,skiasharp,tesseract,openai-gpt4o}. We detail the architecture, token-budgeted batching strategy, and persistence design, then describe a testing plan for grouping quality, time-on-task, and usability. The approach avoids brittle handwriting OCR while preserving instructor control, fairness, and auditability. We conclude with limitations, risks, and suggested future work.
\end{abstract}

@@

\section{Specific Project Goals/Requirements}
The project delivers a practical, instructor-in-the-loop grading system that:
\begin{compactitem}
  \item \textit{Bulk Scans (Filled-form):} PDFs are split by known page counts. Ghostscript rasterizes pages; PdfPig validates page counts; SkiaSharp crops defined regions to PNGs.
  \item \textit{Identity OCR (filled-form only):} Tesseract extracts name/ID from a designated identity region to auto-assign submissions; unresolved items fall back to a manual pick-list. No OCR is performed on answers; grouping uses GPT-4o Vision directly on the cropped images.
  \item \textit{Identity Matching:} (filled-form) identity text is matched to the class roster; (free-form) uploads are automatically tied to the uploader's account.
  \item \textit{Editable AI Assistance:} Vision LLM proposes semantic groups; instructors can merge/split groups, move answers, and apply rubric items per-group.
  \item \textit{Traceability:} All actions and groupings are persisted with timestamps for audit; grades are summarized in an on-screen table for review/copy.
\end{compactitem}

\section{Motivation and Benefits}
\begin{compactitem}
  \item \textbf{Faster feedback:} Instructors grade clusters of similar answers once, reducing turnaround time.
  \item \textbf{Consistency:} Per-group rubric application reduces drift across similar answers and sessions.
  \item \textbf{Lower cognitive load:} The system automates extraction, grouping suggestions, and grade totals; instructors focus on judgment.
  \item \textbf{Reduced brittleness:} Avoiding handwriting OCR on answers eliminates a major failure mode.
  \item \textbf{Privacy-aware:} Only identity crops contain PII; answer crops are devoid of names or IDs.
\end{compactitem}

@@

% -------------------------------------------------
\chapter{Background and Context}\label{chap:bg}
\section{AI in Education}
AI has long promised efficiency gains and personalization in education, from adaptive tutoring to analytics that help instructors intervene earlier. Reviews highlight benefits such as individualized practice, faster feedback, and administrative automation when deployed with appropriate oversight\cite{Alto2023,Chen2020,Luckin2016,Zawacki2019}. Teacher agency, transparency, and contestability remain essential for trust and learning effectiveness.

@@

\section{Bias, Fairness, and Student Perceptions}
Bias can propagate into grading unless monitored and mitigated\cite{Mehrabi2021}. Student acceptance depends on clear processes, contestability, and fairness\cite{Tossell2023}. Effective formative feedback principles—timely, specific, actionable—remain central whether drafted by AI or humans\cite{Nicol2006}.

@@

\paragraph{Commercial paper-exam graders (e.g., \Brand{Gradescope}, \Brand{Crowdmark}).}
\Brand{Gradescope}~\cite{gradescope} supports fixed-template paper exams with region-based workflows and
\enquote{Answer Groups} that let instructors grade clusters of similar responses at once. However, the
grouping method is not publicly documented and is presented at a high level as similarity-based.
\Brand{Crowdmark}~\cite{crowdmark} provides strong scanning workflows (QR-coded booklets, automated
student matching via OCR on cover pages) and can auto-grade multiple choice, but does not claim
semantic grouping of open-ended answers. \textbf{Similarity:} our work also supports fixed templates,
grouping, and rubrics-first grading. \textbf{Difference:} we group from \emph{images only} (no handwriting
OCR of answers), use a token-budgeted vision-LLM pipeline to bound cost/latency, and archive
prompts + model versions for auditability.

\paragraph{OMR/MCQ scanning (e.g., \Brand{Akindi}, \Brand{ZipGrade}).}
\Brand{Akindi}~\cite{akindi} and \Brand{ZipGrade}~\cite{zipgrade} excel at high-throughput multiple-choice grading from
bubble sheets (including mobile scanning) and logistics like sheet sorting. \textbf{Similarity:} we likewise
handle identity intake for large cohorts. \textbf{Difference:} OMR tools target selected-response scoring,
not clustering of free-form handwritten work.

\paragraph{LMS graders (e.g., \Brand{Canvas} SpeedGrader).}
Platform-native graders such as \Brand{Canvas} SpeedGrader~\cite{canvas-speedgrader} offer annotation and rubric
workflows for uploaded files. \textbf{Similarity:} we present rubric-based grading and feedback at scale.
\textbf{Difference:} LMS graders do not automatically cluster semantically similar answers for batch grading.

\paragraph{Autograding/algorithmic assessment (e.g., \Brand{PrairieLearn}, \Brand{M{\"o}bius}, \Brand{CodeRunner}/\Brand{CodeGrade}).}
Systems like \Brand{PrairieLearn}~\cite{prairielearn}, \Brand{M{\"o}bius}~\cite{mobius}, \Brand{CodeRunner}~\cite{coderunner}, and
\Brand{CodeGrade}~\cite{codegrade} autograde parameterized or code questions (randomized variants, unit tests,
CAS checks) with excellent coverage in constrained domains. \textbf{Similarity:} automation reduces
repetitive grader effort. \textbf{Difference:} their strength is \emph{automatic scoring} of structured
responses; they do not aim to \emph{group} heterogeneous, handwritten short answers for a human-in-the-loop
rubric pass.

\paragraph{Positioning.}
Our system is closest in spirit to the \enquote{answer grouping} idea in commercial paper-graders, but
our distinctives are: (1) \textbf{image-only} grouping of handwritten content to avoid brittle OCR,
(2) a \textbf{cost/latency–bounded} vision-LLM pipeline (downscale + tiling + batching), and
(3) an explicitly \textbf{auditable, instructor-in-the-loop} workflow (merge/split/move, neutral labels,
review status), integrated end-to-end with our site.

\section{Takeaways for This Project}
(1) Group-based grading is an effective accelerator; (2) LLMs help when auditable and editable; (3) OCR of handwriting is fragile—visual grouping bypasses failure modes; (4) fairness and oversight practices must be designed-in.

@@

\section{High-Level Architecture}
\begin{figure}[htb]
  \centering
  \resizebox{\textwidth}{!}{\input{images/architecture.tikz}}
  \caption{High-level components and data flow.}
  \label{fig:architecture}
\end{figure}

\section{Sequence of Operations}
\begin{figure}[htb]
  \centering
  \input{images/sequence-autogroup.tikz}
  \caption{Sequence for an auto-grouping job.}
  \label{fig:sequence}
\end{figure}

\begin{compactenum}
  \item Web app posts \texttt{/autogroup} with per-answer image paths and per-question max points.
  \item FastAPI converts PNG$\rightarrow$JPEG (quality 50), computes a token budget at 50\% downscale, and sends images with \texttt{detail:auto}.
  \item GPT-4o Vision proposes clusters; service shapes results (UUIDs, neutral descriptions, collapse tiny groups, add \emph{Ungrouped}).
  \item Results persist to MySQL (one row/question); UI polls \texttt{/status/\{job\_id\}} then renders groups for review and grading.
\end{compactenum}

\section{Instructor-Facing UI Snapshots}
\paragraph{Assignment creation.}
Instructors configure the submission layout (\emph{filled-form} vs.\ \emph{free-form}), directions, and dates. The authoring surface then adapts:

\begin{compactitem}
  \item \textbf{Free-form:} the instructor enters \emph{only} the question labels (e.g., \texttt{Q1}, \texttt{Q2a}) and max points per question. No region editor is shown here because students will upload a PDF and \emph{define their own answer regions} in the next step.
  \item \textbf{Filled-form:} in addition to questions/points, the instructor binds an exam template and draws the \emph{identity} and \emph{answer} regions once. Those regions are then used to crop all submissions automatically.
\end{compactitem}

@@

\paragraph{Course assignments page.}
The course view summarizes assignment state (open/closed, submissions, grading progress) and links to grouping/grading.

@@

\paragraph{Region extraction (shared tool).}
The same cropping widget is used to verify identity/answer regions for filled-form scans and, in free-form flows, to let students mark their own answer areas.

@@

\paragraph{Auto-grouping UI.}
After crops are generated, the grouping page proposes semantic clusters; instructors can merge/split groups, move items, and apply rubric items per group.

@@

\section{Region Extraction and File Lifecycle}
For \textbf{filled-form} assignments, instructors define identity and answer regions once per assignment; the worker then applies those regions to every scanned booklet and enforces stable filenames (e.g., \texttt{Q27a.png}) for reproducibility and idempotent re-uploads. For \textbf{free-form}, students upload a PDF and mark their own answer regions; the saved boxes are used to generate per-question crops for grouping and grading. Debug images are suppressed outside development builds. Re-uploads replace prior crops and metadata to avoid drift.

\section{Identity Assignment}
Filled-form batches use Tesseract to extract roster identifiers from the identity region. Low-confidence matches are flagged for manual resolution with a roster pick-list and thumbnail preview. Free-form uploads are tied to the uploader’s account; no answer OCR is executed.

@@

\section{Data Model}
\begin{table}[htb]
  \centering
  \caption{Key table: \texttt{GroupingResults}.}
  \label{tab:groupingresults}
  \setlength\tabcolsep{6pt}
  \renewcommand{\arraystretch}{1.2}
  \footnotesize
  \begin{tabularx}{\textwidth}{@{}L{4.3cm}L{2.8cm}Y@{}}
    \toprule
    \textbf{Column} & \textbf{Type} & \textbf{Notes} \\
    \midrule
    Id & BIGINT (PK) & Surrogate primary key. \\
    AssignmentId & BIGINT (FK) & Links to the assignment entity. \\
    AssignmentQuestionId & BIGINT (FK) & Equals \texttt{template\_region\_id} sent by client. \\
    GroupData & JSON & Array of groups with files, description, \texttt{is\_correct}, \texttt{points}. \\
    \makecell[l]{CreatedAt\\UpdatedAt} & DATETIME & Audit timestamps (UTC). \\
    \bottomrule
  \end{tabularx}
\end{table}

\section{Prompting \& JSON Schema}
The service uses a system message with practical grouping guidelines (fewer, larger clusters; neutral labels; unreadable/singletons $\rightarrow$ \emph{Ungrouped}). We archive the exact prompts and model/version with the \texttt{job\_id} for reproducibility. The model returns JSON of the following form (abbrev.):
{\small
\begin{verbatim}
{
  "groups": [
    {
      "group_id": "uuid",
      "files": [{ "file_path": "...", "user_id": "...",
                  "template_region_id": "...", "question_label": "...",
                  "content_type": "image/jpeg", "user_name": "..." }],
      "description": "Group 1",
      "is_correct": true | false | null,
      "points": 5.0
    }
  ]
}
\end{verbatim}
}

@@

@@

\paragraph{Filled-form (read-only).}
Students cannot upload; they can download the submitted booklet (when present) and view the grade once posted. The page surfaces \texttt{Directions} and a simple status panel.

@@

\section{Rubrics and Grading UX}
While grading a group, instructors select and apply rubric items; zoom/pan is supported where available. Changing a rubric value propagates to all affected answers. The Question tab surfaces review status: each group displays a status badge (\enquote{Graded} or \enquote{Needs grading}). Instructors can either apply at least one rubric item or, if none are needed, click \texttt{Save All Groups for Question} to mark the question as reviewed.

\section{Security \& Privacy}
Only course roster identifiers and per-answer images are processed; no plaintext student content is transmitted for grouping. Access is limited to authenticated instructor actions. The design aligns with FERPA expectations around access control and least-privilege handling of student records \cite{FERPA1974}.

\section{Threat Model \& Data Handling}
\begin{table}[H]\centering
\caption{Abbreviated threat model and mitigations}\label{tab:threats}
\begin{tabularx}{\textwidth}{@{}p{3.8cm}Xp{4.6cm}@{}}
\toprule
\textbf{Threat} & \textbf{Risk} & \textbf{Mitigation} \\
\midrule
Unauthorized access & Disclosure of student data & Role-based auth; per-course ACLs; audit logs \\
Model data exposure & PII leakage to third party & Only identity crops contain PII; answer crops exclude names \\
Prompt injection & Manipulated grouping suggestions & Server-side prompts; normalize outputs; edit/override tools \\
Data retention & Oversharing over time & Rotation policy; per-course retention config; export/purge \\
\bottomrule
\end{tabularx}
\end{table}

>>>>

% ===================== Packages =====================
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[sc,osf]{mathpazo}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{float} % for [H] placement if desired
\usepackage{tikz} % Diagrams
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,matrix}
\usepackage{siunitx} % Optional but useful
\sisetup{detect-all}
\PassOptionsToPackage{hyphens}{url} % Make long URLs break better in refs (must come before hyperref)
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[nameinlink,capitalize,noabbrev]{cleveref}
\usepackage{bookmark} % better PDF outlines
\usepackage{csquotes} % for \enquote
\usepackage{tabularx} % better tables that auto-wrap
\urlstyle{same}
\usepackage{listings}
\usepackage{lstautogobble} % trims common indent
\usepackage{array}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{subcaption} % defines the subfigure environment
\usepackage{placeins}
\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  breakatwhitespace=false, % Allow breaks anywhere (fixes long strings like connection URLs)
  breakindent=20pt, % Indent wrapped lines for better readability in .cs files
  postbreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\color{gray}\hookrightarrow\space}}, % Subtle continuation arrow for wrapped lines
  columns=fixed, % Changed to fixed for true monospace and even frame lines (prevents uneven extensions)
  keepspaces=true,
  showstringspaces=false,
  frame=single,
  framerule=0.5pt, % Thinner frame rule to reduce visual emphasis on any minor irregularities
  rulecolor=\color{gray}, % Gray frame for subtlety
  framesep=5pt, % Inner padding to keep content from edges
  autogobble=true,
  xleftmargin=5pt, % Slightly wider left margin to prevent hugging
  xrightmargin=5pt, % Matching right margin
  tabsize=4 % Standardize tabs (fixes potential "weird" alignment in .cs code)
}

@@

\begin{abstract}
We present a grading system that accelerates evaluation of open-ended student work across scanned and digital workflows. The system crops answer regions from PDFs, assigns submissions via OCR on identity regions only, and groups answers by visual semantics using a vision LLM. Instructors review and edit groups, apply rubric items once per group, and export grades from an on-screen table. The solution integrates Ghostscript rasterization, PdfPig page orchestration, SkiaSharp region extraction, Tesseract identity OCR, and GPT-4o Vision for grouping\cite{ghostscript,pdfpig,skiasharp,tesseract,openai-gpt4o}. We detail the architecture, token-budgeted batching strategy, and persistence design, then describe testing results for grouping quality, time-on-task, and usability. The approach avoids brittle handwriting OCR while preserving instructor control, fairness, and auditability.
\end{abstract}

@@

\section{Specific Project Goals/Requirements}
This project delivers a practical, instructor-in-the-loop grading system that:
\begin{compactitem}
  \item \textit{Bulk Scans (Filled-form):} PDFs are split by known page counts. Ghostscript rasterizes pages; PdfPig validates page counts; SkiaSharp crops defined regions to PNGs.
  \item \textit{Identity OCR (filled-form only):} Tesseract extracts name/ID from a designated identity region to auto-assign submissions; unresolved items fall back to a manual pick-list. No OCR is performed on answers; grouping uses GPT-4o Vision directly on the cropped images.
  \item \textit{Identity Matching:} (filled-form) identity text is matched to the class roster; (free-form) uploads are automatically tied to the uploader's account.
  \item \textit{Editable AI Assistance:} Vision LLM proposes semantic groups; instructors can merge/split groups, move answers, and apply rubric items per-group.
  \item \textit{Traceability:} All actions and groupings are persisted with timestamps for audit; grades are summarized in an on-screen table for review/copy.
\end{compactitem}

\section{Motivation and Benefits}
\begin{compactitem}
  \item \textbf{Faster feedback:} Instructors grade clusters of similar answers once, reducing turnaround time.
  \item \textbf{Consistency:} Per-group rubric application reduces drift across similar answers and sessions.
  \item \textbf{Lower cognitive load:} The system automates extraction, grouping suggestions, and grade totals; instructors focus on assessment.
  \item \textbf{Reduced brittleness:} Avoiding handwriting OCR on answers eliminates a major failure mode.
  \item \textbf{Privacy-aware:} Only identity crops contain PII; answer crops are devoid of names or IDs.
\end{compactitem}

@@

% -------------------------------------------------
\chapter{Background and Context}\label{chap:bg}
\section{AI in Education}
AI has long promised efficiency gains and personalization in education, from adaptive tutoring to analytics that help instructors intervene earlier. Reviews highlight benefits such as individualized practice, faster feedback, and administrative automation when deployed with appropriate oversight\cite{Alto2023,Chen2020,Luckin2016,Zawacki2019}. Ensuring teachers can review and revise AI-generated grades—and retain final say to confirm fairness—is essential for trust and real learning gains.

@@

\section{Bias, Fairness, and Student Perceptions}
Bias can propagate into grading unless monitored and mitigated\cite{Mehrabi2021}. Student acceptance depends on clear processes, and fairness\cite{Tossell2023}. Effective formative feedback principles—timely, specific, actionable—remain central whether drafted by AI or humans\cite{Nicol2006}.

@@

\subsection{Commercial paper-exam graders (e.g., \Brand{Gradescope}, \Brand{Crowdmark})}
\Brand{Gradescope}~\cite{gradescope} supports fixed-template paper exams with region-based workflows and
\enquote{Answer Groups} that let instructors grade clusters of similar responses at once. However, the
grouping method is not publicly documented and is presented at a high level as similarity-based.
\Brand{Crowdmark}~\cite{crowdmark} provides strong scanning workflows (QR-coded booklets, automated
student matching via OCR on cover pages) and can auto-grade multiple choice, but does not claim
semantic grouping of open-ended answers. \textbf{Similarity:} our work also supports fixed templates,
grouping, and rubrics-first grading. \textbf{Difference:} we group from \emph{images only} (no handwriting
OCR of answers), use a token-budgeted vision-LLM pipeline to bound cost/latency, and archive
prompts + model versions for auditability.

\subsection{OMR/MCQ scanning (e.g., \Brand{Akindi}, \Brand{ZipGrade})}
\Brand{Akindi}~\cite{akindi} and \Brand{ZipGrade}~\cite{zipgrade} excel at high-throughput multiple-choice grading from
bubble sheets (including mobile scanning) and logistics like sheet sorting. \textbf{Similarity:} we likewise
handle identity intake for large cohorts. \textbf{Difference:} OMR tools target selected-response scoring,
not clustering of free-form handwritten work.

\subsection{LMS graders (e.g., \Brand{Canvas} SpeedGrader)}
Platform-native graders such as \Brand{Canvas} SpeedGrader~\cite{canvas-speedgrader} offer annotation and rubric
workflows for uploaded files. \textbf{Similarity:} we present rubric-based grading and feedback at scale.
\textbf{Difference:} LMS graders do not automatically cluster semantically similar answers for batch grading.

\subsection{Autograding/algorithmic assessment (e.g., \Brand{PrairieLearn}, \Brand{M{\"o}bius}, \Brand{CodeRunner}/\Brand{CodeGrade})}
Systems like \Brand{PrairieLearn}~\cite{prairielearn}, \Brand{M{\"o}bius}~\cite{mobius}, \Brand{CodeRunner}~\cite{coderunner}, and
\Brand{CodeGrade}~\cite{codegrade} autograde parameterized or code questions (randomized variants, unit tests,
CAS checks) with excellent coverage in constrained domains. \textbf{Similarity:} automation reduces
repetitive grader effort. \textbf{Difference:} their strength is \emph{automatic scoring} of structured
responses; they do not aim to \emph{group} heterogeneous, handwritten short answers for a human-in-the-loop
rubric pass.

\subsection{Positioning}
Our system is closest in spirit to the \enquote{answer grouping} idea in commercial paper-graders, but
our distinctives are: (1) \textbf{image-only} grouping of handwritten content to avoid brittle OCR,
(2) a \textbf{cost/latency–bounded} vision-LLM pipeline (downscale + tiling + batching), and
(3) an explicitly \textbf{auditable, instructor-in-the-loop} workflow (merge/split/move, neutral labels,
review status), integrated end-to-end with our site.

\section{Conclusions from Research}
(1) Group-based grading is an effective accelerator; (2) LLMs help when auditable and editable; (3) OCR of handwriting is fragile—visual grouping bypasses failure modes; (4) fairness and oversight practices must be designed-in.

@@

\section{High-Level Architecture}
At a glance, the platform is organized into three dashed \emph{zones} that separate concerns and scaling boundaries: the \emph{Application Server} (Razor Pages web app and extraction worker), the \emph{AI Service Layer} (FastAPI microservice and the GPT-4o Vision endpoint), and \emph{Data \& Storage} (MySQL and the file store). Instructors interact via a browser over HTTPS with the Razor Pages app, which handles authentication, rubric management, and the grading UI. The app invokes an extraction worker that orchestrates Ghostscript, PdfPig, and SkiaSharp to render pages and cut out per-answer crops; Tesseract OCR is used narrowly to read identifying fields (e.g., student ID) and is deliberately decoupled from semantic grouping. Crops and downstream exports are written to the file store while the web app serves these images directly to the UI.

Auto-grouping requests are queued from the web app to a Python FastAPI service (\texttt{POST /autogroup}). The service packages each question’s crops, converts them to compact JPEGs, and sends them with prompt instructions to GPT-4o Vision. Proposed clusters are normalized server-side (stable UUIDs, neutral names, small-group collapse) and persisted to MySQL. The UI polls a job status endpoint and, when complete, renders groups for human review, edits, and grading. This arrangement keeps the web tier largely stateless, allows the extraction worker and AI service to scale independently, and localizes persistent state to MySQL and the file store. The overall topology and data flow are depicted in Figure~\ref{fig:architecture}.

\begin{figure}[htb]
  \centering
  \resizebox{\textwidth}{!}{\input{images/architecture.tikz}}
  \caption{High-level components and data flow.}
  \label{fig:architecture}
\end{figure}

\section{Sequence of Operations}
An instructor initiates grouping from the web UI by selecting a question and confirming its configuration (maximum points, rubric, and the set of per-answer image paths). The browser submits this intent to the web app, which records a new job and enqueues a request to the FastAPI microservice at \texttt{/autogroup}. In parallel or beforehand (depending on question state), the extraction worker renders the relevant PDF pages with Ghostscript, enumerates answer regions via PdfPig, and uses SkiaSharp to crop each region to PNG. Lightweight OCR with Tesseract runs only on identifying fields (e.g., cover-page name/ID boxes) to support later reconciliation; the content of answers themselves is not OCRed for grouping. All crops are written to the file store under stable, human-inspectable paths, and the web app exposes read-only URLs so the UI can preview exactly what will be grouped.

Upon receiving an \texttt{/autogroup} task, the FastAPI service prepares the model payload. Each PNG crop is converted to JPEG at a target quality of 50 to reduce bandwidth and context size while preserving legibility for short answers. The service computes a token budget using a 50\% downscale heuristic that caps the number of pixels sent per image; if the source crop exceeds that budget, it is downsampled while maintaining aspect ratio. The prompt attaches each image with \texttt{detail:auto} and instructs the model to propose clusters of semantically similar answers. The prompt also requests that low-confidence or outlier responses be flagged for an \emph{Ungrouped} bucket.

GPT-4o Vision returns candidate clusters that the service further \emph{shapes} before persistence. Each cluster receives a stable UUID so subsequent UI edits (rename, merge, split) can be tracked independently of order. Neutral descriptions are standardized (e.g., using a canonical exemplar from the cluster) and very small clusters are folded into \emph{Ungrouped} based on a configurable minimum size. Optionally, if embeddings are available, near-duplicate microclusters are merged with a conservative similarity threshold to avoid over-fragmentation. The shaped result---including cluster membership, labels, and an audit of any collapsed/ungrouped items---is written to MySQL as one row per question with a foreign key to the job.

While the job runs, the UI polls \texttt{/status/\{job\_id\}} with backoff to avoid excessive load. When the job is \texttt{Complete}, the browser fetches the grouped payload and renders cluster cards with thumbnails backed by the file store. Instructors may rename clusters, reassign individual answers, or merge/split clusters as needed; those edits are persisted incrementally. When grading begins, the UI ties rubric criteria and maximum points to each cluster, enabling a single grading action to fan out to all member answers. Final scores are written through to the \texttt{GroupingResults} and \texttt{GroupScore} tables. The complete message choreography for this workflow is shown in the sequence diagram in Figure~\ref{fig:sequence}.

\begin{figure}[htb]
  \centering
  \input{images/sequence-autogroup.tikz}
  \caption{Sequence for an auto-grouping job.}
  \label{fig:sequence}
\end{figure}

\section{Instructor-Facing UI Snapshots}
This section highlights the key instructor workflows with inline references to the corresponding UI figures: assignment creation (Figure~\ref{fig:assignment-creation}), the course assignments overview (Figure~\ref{fig:course-page}), the shared region-cropping tool (Figure~\ref{fig:ui-crops}), and the auto-grouping interface (Figure~\ref{fig:ui-grouping}).

\paragraph{Assignment creation.}
Instructors configure the submission layout (\emph{filled-form} vs.\ \emph{free-form}), directions, and dates. As shown in Figure~\ref{fig:assignment-creation}, the authoring adapts to the chosen layout: 
\begin{compactitem}
  \item \textbf{Free-form:} the instructor enters \emph{only} the question labels (e.g., \texttt{Q1}, \texttt{Q2a}) and max points per question. No region editor is shown here because students will upload a PDF and \emph{define their own answer regions} in the next step (see also the cropping widget in Figure~\ref{fig:ui-crops}).
  \item \textbf{Filled-form:} in addition to questions/points, the instructor binds an exam template and draws the \emph{identity} and \emph{answer} regions once. Those regions are then used to crop all submissions automatically (the same widget used here is illustrated in Figure~\ref{fig:ui-crops}).
\end{compactitem}

@@

\paragraph{Course assignments page.}
The course view summarizes assignment state (open/closed, submissions, grading progress) and links to grouping/grading. Figure~\ref{fig:course-page} shows status indicators and quick actions that guide instructors into grouping (Figure~\ref{fig:ui-grouping}) or back to the region editor (Figure~\ref{fig:ui-crops}) if setup needs adjustment.

@@

\paragraph{Region extraction (shared tool).}
The same cropping widget is used to verify identity/answer regions for filled-form scans and, in free-form flows, to let students mark their own answer areas. Figure~\ref{fig:ui-crops} depicts the shared tool; the resulting crops flow directly into the grouping experience shown in Figure~\ref{fig:ui-grouping}.

@@

\paragraph{Auto-grouping UI.}
After crops are generated, the grouping page proposes semantic clusters; instructors can merge/split groups, move items, and apply rubric items per group. Figure~\ref{fig:ui-grouping} shows the proposed clusters and edit tools; rubric-first grading executed here propagates scores to all members of a cluster.

@@

\section{Region Extraction and File Lifecycle}
For \textbf{filled-form} assignments, instructors define identity and answer regions once per assignment using the cropping widget (Figure~\ref{fig:ui-crops}); the worker then applies those regions to every scanned booklet and enforces stable filenames (e.g., \texttt{Q27a.png}) for reproducibility and idempotent re-uploads. For \textbf{free-form} assignments, students upload a PDF and mark their own answer regions in the same tool (Figure~\ref{fig:ui-crops}); the saved boxes are used to generate per-question crops that populate the grouping interface (Figure~\ref{fig:ui-grouping}) for review and rubric-first grading. Debug images are suppressed outside development builds. Re-uploads replace prior crops and metadata to avoid stale data, and the updated crops reappear in the grouping view (Figure~\ref{fig:ui-grouping}) so that any instructor edits remain aligned with the latest artifacts.

\section{Identity Assignment}
Filled-form batches use Tesseract~\cite{tesseract} to extract roster identifiers from the pre-defined \emph{identity} region (see Figure~\ref{fig:ui-crops}). Low-confidence or malformed results are flagged for manual resolution with a roster pick-list and a thumbnail preview. Free-form uploads are tied to the uploader’s account; therefore no OCR is needed. Identity OCR is used only for roster linkage—semantic grouping operates directly on cropped answer images (Section~\ref{chap:solution}, Figures~\ref{fig:architecture} and~\ref{fig:sequence}).

@@

\section{Data Model}
Auto-grouping results are persisted as a \emph{single row per question}, providing a denormalized snapshot that the UI can load quickly after the workflow in Figure~\ref{fig:sequence} and during review in Figure~\ref{fig:ui-grouping}. As summarized in Table~\ref{tab:groupingresults}, the JSON payload \texttt{GroupData} holds the proposed/edited clusters (neutral labels, membership, per-group points/flags), while \texttt{ScoringScheme} and the boolean guards (\texttt{AllowBelowZero}, \texttt{AllowAboveMax}) control rubric arithmetic. \texttt{CreatedAt}/\texttt{UpdatedAt} provide auditability. In practice, there is at most one record per \texttt{(AssignmentId, AssignmentQuestionId)}; edits typically mutate \texttt{GroupData} and bump \texttt{UpdatedAt}.

\begin{table}[H]
  \centering
  \caption{Key table: \texttt{GroupingResults}.}
  \label{tab:groupingresults}
  \setlength\tabcolsep{6pt}
  \renewcommand{\arraystretch}{1.2}
  \footnotesize
  \begin{tabularx}{\textwidth}{@{}L{4.3cm}L{2.8cm}Y@{}}
    \toprule
    \textbf{Column} & \textbf{Type} & \textbf{Notes} \\
    \midrule
    Id & INT (PK) & Surrogate primary key. \\
    AssignmentId & INT (FK) & Links to the assignment entity. \\
    AssignmentQuestionId & INT (FK) & Equals \texttt{template\_region\_id} sent by client. \\
    GroupData & JSON & Array of groups with files, description, \texttt{is\_correct}, \texttt{points}. \\
    ScoringScheme & VARCHAR(32) & Default \texttt{"Negative"}. \\
    AllowBelowZero & TINYINT(1) & Boolean. \\
    AllowAboveMax & TINYINT(1) & Boolean. \\
    CreatedAt & DATETIME & UTC created timestamp. \\
    UpdatedAt & DATETIME NULL & UTC last update; nullable. \\
    \bottomrule
  \end{tabularx}
\end{table}

\paragraph{Prompting \& JSON schema.}
The service uses concise grouping rules (favor fewer, larger clusters; neutral labels; unreadable/singletons $\rightarrow$ \emph{Ungrouped}). Prompts and model/version are archived with each \texttt{job\_id}. The model returns JSON in the following abbreviated form:
\begin{lstlisting}[language=json]
{
  "groups": [
    {
      "group_id": "uuid",
      "files": [
        {
          "file_path": "...",
          "user_id": "...",
          "template_region_id": "...",
          "question_label": "...",
          "content_type": "image/jpeg",
          "user_name": "..."
        }
      ],
      "description": "Group 1",
      "is_correct": true | false | null,
      "points": 5.0
    }
  ]
}
\end{lstlisting}

@@

\section{Service Isolation \& Network Security}
In production, the FastAPI auto-grouping service runs inside the same Docker Compose stack on a private bridge network and is not published to the host (container uses \texttt{expose} only—no \texttt{ports} mapping). As a result, \texttt{/autogroup} and related endpoints are reachable only from the web app over internal service DNS (e.g., \texttt{http://autogroup:8000/...}); the browser never talks to the service directly. Compose-level network isolation and host firewall rules prevent external ingress to the service container.

@@

\paragraph{Filled-form (read-only).}
Students cannot upload; they can download the submitted booklet (when present) and view the grade once posted. The page shows \texttt{Directions} and a simple status panel.

@@

\section{Rubrics and Grading UX}
While grading a group, instructors select and apply rubric items; zoom/pan is supported where available. Changing a rubric value propagates to all affected answers. The Question tab shows review status: each group displays a status badge (\enquote{Graded} or \enquote{Needs grading}). Instructors can either apply at least one rubric item or, if none are needed, click \texttt{Save All Groups for Question} to mark the question as reviewed.

\section{Security \& Privacy}
We minimize student-data exposure by limiting OCR strictly to identity regions (therefore no OCR is needed on answers), sending only per-answer image crops—without student names—to the vision model for grouping, and confining operations to authenticated instructor actions with role-appropriate permissions. Only course roster identifiers and per-answer images are processed; no plaintext student content is transmitted for grouping. All reads and writes are logged for auditability, and the design follows least-privilege access control consistent with FERPA expectations~\cite{FERPA1974}. To safeguard fairness, we monitor for potential bias by auditing cluster assignments across demographic-neutral cohorts and preserve human authority through full instructor override mechanisms for grouping and grading.

\section{Threat Model \& Data Handling}
This section details how identities, images, and grouping results are protected throughout the workflow in Figure~\ref{fig:sequence} and during review in Figure~\ref{fig:ui-grouping}. For a concise summary, see Table~\ref{tab:threats} at the end of this section.
